\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-10-07 appendices.tex
went through proof of Theorem 5

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$, we first note that without loss of generality, each message $B_i$ can is of the form
\begin{equation}
\label{eq:general_messages}
B_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \R$ is a Borel measurable set. 
%Indeed, any measurable function $(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
Consider the conditional distribution $P({B^n|\theta})$ of $B^n$ given $\theta$. We have 
\begin{align}
P\left( B^n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B^{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(B_i =1 | \theta, B^{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$, so that the Fisher information of $B^n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(B^n) = \sum_{i=1}^n I_\theta (B_i|B^{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B^{i-1})$ is the Fisher information of the distribution of $B_i$ given $B^{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (B_i|B^{i-1}) \leq 4f^2(0)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B^n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}.
\end{align*}
\QEDA

@@

\begin{lem} \label{lem:bound_intervals_delta}
Let $f(x)$ be log-concave, symmetric, and differentiable density function. Let $\delta\geq 0$. Assume that the origin $x=0$ uniquely maximizes the function
\[
\eta_\delta(x) \triangleq  \eta^{1+\delta}(x)/f^\delta(x) = \frac{  \left( f(x) \right)^{2+\delta}}{\left(F(x)(1-F(x))\right)^{1+\delta}}
\]
and that $\eta_\delta(x)$ is non-increasing in $|x|$. For any $x_1 \ge \ldots \ge x_n \in \R^n$,
\begin{equation}
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left| \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{k=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} } 
\leq  \max_{i} \eta_\delta(x_i).
\label{eq:lem_bound_intervals_delta}
\end{equation}
\end{lem}
In particular, 
\[
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left| \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} } 
\leq \eta_\delta(0) = 4^{1+\delta} f^{2+\delta}(0).
\]

@@

The log probability mass distribution of $B^n=(B_1,\ldots,B_n)$ is given by

@@

& \log \Prob_\theta(b^n) = \sum_{i=1}^n \left( \frac{b_i+1}{2} \log  \Prob(X_i\in A_i) + 
\frac{1-b_i}{2} \log \Prob(X_i\in A_i) \right), \quad b^n \in \{-1,1\}^n. 

@@

& \log \frac{ \Prob_{\theta + \frac{h}{\sqrt{n}}}(b^n)} { \Prob_\theta (b^n) } = \sum_{i=1}^n    \frac{b_i+1}{2} 

@@

A_i = \bigcup_{k=1}^{K_i} \left(t_{i,k},t_{i,k+1} \right),

@@

\Prob_{\theta}(X_i \in A_i) = \sum_{k=1}^{K_i} (-1)^k F(x_{i,k}-\theta).

@@

In particular, since $f$ is differentible, $\Prob_{\theta}(X_i \in A_i)$ is twice differentiable, and we may write

@@

& \log \frac{ \Prob_{\theta + \frac{h}{\sqrt{n}}}(b^n)} { \Prob_\theta (b^n) }   = \frac{h}{\sqrt{n}} \sum_{i=1}^n  

@@

Next, we claim that there exits $\delta>0$ and $K>0$, that are independent of $n$, such that

@@

\frac{1}{n} \sum_{i=1}^n \ex{ |U_i|^{2+\delta}}  < K \label{eq:Lyaponov_num}

@@

  { (\Prob_{\theta}(X_i \in A_i))^{1+\delta} (1-\Prob_{\theta}(X_i \in A_i))^{1+\delta}}, 

@@

where the last transition is because 
 \left( (1-\Prob_{\theta}(X_i \in A_i))^{1+\delta} + (\Prob_{\theta}(X_i \in A_i))^{1+\delta}  \right) \leq 1. 
We now use the fact that each $A_i$ is a finite union of intervals. Under the assumption that there exists $\delta>0$ such that $\eta^{1+\delta}(x)/f^\delta(x)$ is maximized by the origin, we get
\eta^{1+\delta}(x)/f^\delta(x) = \frac{(f(x))^{2+\delta}}{\left(F(x)(1-F(x))\right)^{1+\delta}} \leq 4^{1+\delta} (f(0))^{2+\delta}. 
Lemma~\ref{lem:bound_intervals_delta} implies

@@

and thus the numerator of \eqref{eq:Lyaponov}, as well as the entire expression, goes to zero. From Lyaponov's central limit theorem we conclude that 

@@

Let $\Xi$ be the set of points $\theta \in \Theta$ for which $\kappa(\theta) = \eta(0)$. 
%
Since $B_1,B_2,\ldots$ satisfy the conditions in Theorem~\ref{thm:LAN1}, $\theta$ is in $\Xi$ if and only if
\begin{equation}
\label{eq:non_existence_proof}
\lim_{n\to \infty} L_n(A_1,\ldots,A_n;\theta) = \eta(0). 
\end{equation}
By assumption, we have $B_i^{-1} = A_i$ where $A_i$ can be expressed as
\[
A_i = \cup_{i=1}^K (a_{i,k},b_{i,k}), 
\]
where $a_{i,1} \leq b_{i,1} \leq \ldots \leq a_{i,K}, b_{i,K}$, and $a_{i,1}$ and $b_{i,K}$ may take the values $-\infty$ and $\infty$, respectively. Denote
\[
\mathcal B_i = \cup_{k=1}^{K}\{a_{i,k},b_{i,k}\}.
\]
For any $\theta$ and $\epsilon>0$, denote 
\[
S_n(\theta, \epsilon) \triangleq \left\{ i\leq n \,:\, (\theta-\epsilon,\theta+\epsilon) \cap \mathcal B_i \neq \emptyset \right\}
\]
In words, $S_n$ contains all integers smaller than $n$ in which an $\epsilon$-ball around $\theta$ contains an endpoint of one of the intervals consisting $A_i$. 
%
We now claim that 
if $\theta \in \Xi$ then $\card(S_n(\theta, \epsilon))/n \to 1$. Indeed, for such $\theta$ we have
\begin{align}
& L_n(A_1,\ldots,A_n; \theta) \nonumber \\
& = \frac{1}{n} \sum_{i \in S_n(\epsilon,\theta)}  
\frac{ \left(\sum_{k=1}^{K}  f(\theta - b_{i,k})- f(\theta - a_{i,k}) \right)^2}{ \sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right) \left(1-\sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right)\right)} \nonumber \\
& 
+ \frac{1}{n}\sum_{i \notin S_n(\epsilon,\theta) } \frac{ \left(\sum_{k=1}^{K}  f(b_{i,k}-\theta) - f(a_{i,k}-\theta) \right)^2} { \sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right) \left(1-\sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right)\right)} \nonumber \\
& \overset{(a)}{\leq}
\frac{\card\left(S_n(\theta,\epsilon)\right)}{n} \eta(0) + \frac{n-\card\left(S_n(\theta,\epsilon) \right) }{n} \eta(\epsilon) 
 \label{eq:non_existence_proof1}
\end{align}
where $(a)$ follows from Lemma~\ref{lem:bound_intervals_delta} with $\delta =0$, and the fact that for $i \in S_n(\theta, \epsilon)$, 
\[
\max\left\{ \max_k \eta(b_{i,k}-\theta) , \max_k \eta(a_{i,k}-\theta)  \right\} \leq \eta(\epsilon) < \eta(0). 
\]
Unless  $\card \left(S_n(\theta, \epsilon) \right)/n \to 1$, we get that \eqref{eq:non_existence_proof1}, hence $L_n(A_1,\ldots,A_n ; \theta)$, are bounded from above by a constant that is smaller then $\eta(0)$ in contradiction to the fact that $\theta \in \Xi$. \par
For $k\in \N$, assume by contradiction that there exists $N \geq 2K + 1$ distinct elements
$\theta_1,\ldots,\theta_N \in \Xi$. Since each $A_i$ consists of at most $K$ intervals, we have that 
\begin{equation}
\label{eq:few_optimality_points_proof}
\card (\cup_{i=1}^n \mathcal B_i) \leq 2 n K. 
\end{equation}
Fix $\epsilon>0$ such that 
\[
\epsilon < \frac{1}{2}\min_{i\neq j} |\theta_i - \theta_j|. 
\]
Since for each $\theta \in \Theta$ we have $S_n(\theta, \epsilon) \to 1$, there exists $n$ large enough such that 
\[
\card \left(S_n(\theta_i, \epsilon) \right) \geq n \left(1-\frac{1}{2N} \right)
\]
for all $i=1,\ldots,N$. However, $S_n(\theta_1,\epsilon), \ldots S_n(\theta_N,\epsilon)$ are disjoint, so the cardinallity of their union is at least $n\left(1-\frac{1}{2N} \right)N$ which is grater than $2nK + n/2$ in contradiction to \eqref{eq:few_optimality_points_proof}.

>>>>

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$, we first note that without loss of generality, each message $B_i$ can is of the form
\begin{equation}
\label{eq:general_messages}
B_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \R$ is a Borel measurable set. 
%Indeed, any measurable function $(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
Consider the conditional distribution $P({B_1,\ldots,B_n|\theta})$ of $(B_1,\ldots,B_n)$ given $\theta$. We have 
\begin{align}
P\left( B_1,\ldots,B_n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B_1,\ldots,B_{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(B_i =1 | \theta, B_1,\ldots,B_{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$, so that the Fisher information of $B_1,\ldots,B_n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(B_1,\ldots,B_N) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the distribution of $B_i$ given $B_1,\ldots,B_{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (B_i|B_{i-1},\ldots,B_1) \leq 4f^2(0)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}.
\end{align*}
\QEDA

@@

\begin{lem} \label{lem:bound_intervals_delta}
Let $f(x)$ be log-concave, symmetric, and differentiable density function. Let $\delta\geq 0$. Assume that the function
\[
\eta_\delta(x) \triangleq  \eta^{1+\delta}(x)/f^\delta(x) = \frac{  \left( f(x) \right)^{2+\delta}}{\left(F(x)(1-F(x))\right)^{1+\delta}}
\]
is non-increasing in $|x|$. For any $x_1 \ge \ldots \ge x_n \in \R^n$,
\begin{equation}
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left| \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{k=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} } 
\leq  \max_{i} \eta_\delta(x_i).
\label{eq:lem_bound_intervals_delta}
\end{equation}
\end{lem}
In particular, 
\[
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left| \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} } 
\leq \eta_\delta(0) = 4^{1+\delta} f^{2+\delta}(0).
\]

@@

The log-probability mass distribution of $(B_1,\ldots,B_n)$ is given by

@@

& \log \Prob_\theta(b_1,\ldots,b_n) = \sum_{i=1}^n \left( \frac{b_i+1}{2} \log  \Prob(X_i\in A_i) + 
\frac{1-b_i}{2} \log \Prob(X_i\in A_i) \right), \quad b_i \in \{-1,1\}, \quad i=1,\ldots,n. 

@@

& \log \frac{ \Prob_{\theta + \frac{h}{\sqrt{n}}}(b_1,\ldots,b_n)} { \Prob_\theta (b_1,\ldots,b_n) } = \sum_{i=1}^n    \frac{b_i+1}{2} 

@@

A_i = \bigcup_{k=1}^{K_i} \left(t_{i,2k-1},t_{i,2k} \right),

@@

\Prob_{\theta}(X_i \in A_i) = \sum_{k=1}^{K_i} (-1)^k F(t_{i,k}-\theta).

@@

In particular, since $f$ is differentible in $\theta$, $\Prob_{\theta}(X_i \in A_i)$ is twice differentiable, and we may write

@@

& \log \frac{ \Prob_{\theta + \frac{h}{\sqrt{n}}}(b_1,\ldots,b_n)} { \Prob_\theta (b_1,\ldots,b_n) }   = \frac{h}{\sqrt{n}} \sum_{i=1}^n  

@@

Next, we claim that there exits $\delta>0$ and $K(\delta)>0$, that are independent of $n$, such that

@@

\frac{1}{n} \sum_{i=1}^n \ex{ |U_i|^{2+\delta}}  < K(\delta) \label{eq:Lyaponov_num}

@@

  { (\Prob_{\theta}(X_i \in A_i))^{1+\delta} (1-\Prob_{\theta}(X_i \in A_i))^{1+\delta}}. 

@@

%where the last transition is because 
%\[
% \left( (1-\Prob_{\theta}(X_i \in A_i))^{1+\delta} + (\Prob_{\theta}(X_i \in A_i))^{1+\delta}  \right) \leq 1. 
%\]
We now use the fact that each $A_i$ is a finite union of intervals. Under the assumption that there exists $\delta>0$ such that $\eta^{1+\delta}(x)/f^\delta(x)$ is uniquely maximized by the origin and is non-increasing in $|x|$, Lemma~\ref{lem:bound_intervals_delta} implies

@@

and thus the numerator of \eqref{eq:Lyaponov}, as well as the entire expression, goes to zero. From Lyaponov's central limit theorem we obtain 

@@

Let $\Xi$ be the set of points $\theta \in \Theta$ for which $\kappa(\theta) = \eta(0)$. 
%
Since $B_1,B_2,\ldots$ satisfy the conditions in Theorem~\ref{thm:LAN1}, $\theta$ is in $\Xi$ if and only if
\begin{equation}
\label{eq:non_existence_proof}
\lim_{n\to \infty} L_n(A_1,\ldots,A_n;\theta) = \eta(0). 
\end{equation}
By assumption, we have $B_i^{-1} = A_i$ where $A_i$ can be expressed as
\[
A_i = \cup_{i=1}^K (a_{i,k},b_{i,k}), 
\]
where $a_{i,1} \leq b_{i,1} \leq \ldots \leq a_{i,K}, b_{i,K}$, and $a_{i,1}$ and $b_{i,K}$ may take the values $-\infty$ and $\infty$, respectively. Denote
\[
C_i = \cup_{k=1}^{K}\{a_{i,k},b_{i,k}\}.
\]
For any $\theta$ and $\epsilon>0$, denote 
\[
S_n(\theta, \epsilon) \triangleq \left\{ i\leq n \,:\, (\theta-\epsilon,\theta+\epsilon) \cap C_i \neq \emptyset \right\}
\]
In words, $S_n$ contains all integers smaller than $n$ in which an $\epsilon$-ball around $\theta$ contains an endpoint of one of the intervals consisting $A_i$. 
%
We now claim that 
if $\theta \in \Xi$ then $\card(S_n(\theta, \epsilon))/n \to 1$. Indeed, for such $\theta$ we have
\begin{align}
& L_n(A_1,\ldots,A_n; \theta) \nonumber \\
& = \frac{1}{n} \sum_{i \in S_n(\epsilon,\theta)}  
\frac{ \left(\sum_{k=1}^{K}  f(\theta - b_{i,k})- f(\theta - a_{i,k}) \right)^2}{ \sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right) \left(1-\sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right)\right)} \nonumber \\
& 
+ \frac{1}{n}\sum_{i \notin S_n(\epsilon,\theta) } \frac{ \left(\sum_{k=1}^{K}  f(b_{i,k}-\theta) - f(a_{i,k}-\theta) \right)^2} { \sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right) \left(1-\sum_{k=1}^{K} \left( F(\theta - b_{i,k})- F(\theta - a_{i,k}) \right)\right)} \nonumber \\
& \overset{(a)}{\leq}
\frac{\card\left(S_n(\theta,\epsilon)\right)}{n} \eta(0) + \frac{n-\card\left(S_n(\theta,\epsilon) \right) }{n} \eta(\epsilon) 
 \label{eq:non_existence_proof1}
\end{align}
where $(a)$ follows from Lemma~\ref{lem:bound_intervals_delta} with $\delta =0$, and the fact that for $i \in S_n(\theta, \epsilon)$, 
\[
\max\left\{ \max_k \eta(b_{i,k}-\theta) , \max_k \eta(a_{i,k}-\theta)  \right\} \leq \eta(\epsilon) < \eta(0). 
\]
Unless  $\card \left(S_n(\theta, \epsilon) \right)/n \to 1$, we get that \eqref{eq:non_existence_proof1}, hence $L_n(A_1,\ldots,A_n ; \theta)$, are bounded from above by a constant that is smaller then $\eta(0)$ in contradiction to the fact that $\theta \in \Xi$. \par
For $k\in \N$, assume by contradiction that there exists $N \geq 2K + 1$ distinct elements
$\theta_1,\ldots,\theta_N \in \Xi$. Since each $A_i$ consists of at most $K$ intervals, we have that 
\begin{equation}
\label{eq:few_optimality_points_proof}
\card (\cup_{i=1}^n \mathcal B_i) \leq 2 n K. 
\end{equation}
Fix $\epsilon>0$ such that 
\[
\epsilon < \frac{1}{2}\min_{i\neq j} |\theta_i - \theta_j|. 
\]
Since for each $\theta \in \Theta$ we have $S_n(\theta, \epsilon) \to 1$, there exists $n$ large enough such that 
\[
\card \left(S_n(\theta_i, \epsilon) \right) \geq n \left(1-\frac{1}{2N} \right)
\]
for all $i=1,\ldots,N$. However, $S_n(\theta_1,\epsilon), \ldots S_n(\theta_N,\epsilon)$ are disjoint, so the cardinallity of their union is at least $n\left(1-\frac{1}{2N} \right)N$ which is grater than $2nK + n/2$ in contradiction to \eqref{eq:few_optimality_points_proof}.
